The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-12-22 paper/main.tex
update paper with q^{2/3} results

Using Eq.~\ref{eq:general_Gamma_a}, we can show that when $Z_0=Z_1=Z_2$,
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma\a(r\a,M\a)}{\Gamma_0(r,M)} 
\right|_{r\a\rightarrow r,\ M\a\rightarrow M} \\
&=
\notag
\frac{1}{1+\mu}
\left[1 + \frac{1}{\mathcal{N}_q} \frac{{\rm BF}}{1-{\rm BF}}\times 
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+4}{2}} +
\right.
\right. \\
&\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
\left.\left.
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
(1+q^\alpha)^{\frac{3}{2}}
(1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\label{eq:powerlaw_vary_binary}
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\gamma=0$, $\delta=-2.92$, the 
summed integrals in Eq.~\ref{eq:powerlaw_vary_binary} give $(\ldots)\approx 
1.41425$. %171211_model_integrals.nb
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma_0 = 1.014$; the
apparent rate density is higher than the rate density around singles by 1.4\%.
To summarize, this model suggests that for a power-law radius distribution 
($2r_\oplus \lesssim r\a \lesssim 17r_\oplus$) in which there are the same 
number of planets orbiting singles, primaries, and secondaries, binarity 
influences apparent planet occurrence rates around Sun-like stars at the 
$\sim$few percent level.

@@

At apparent radii $r\a > 2r_\oplus$, the results of this model are the same as 
those from Sec.~\ref{sub:powerlaw_varying_binaries}.
For $r\a < 2r_\oplus$, the equations are quite tedious, though still tractable.
To compute $\Gamma\a(r\a)$, we simply insert 
Eq.~\ref{eq:model3_radius_distribution} into Eq.~\ref{eq:general_Gamma_a}, and 
integrate using a computer program.
We refer the interested reader to our online 
implementation\footnote{\url{https://github.com/lgbouma/binary_biases}}.
The output is validated against analytic predictions in the $r\a > 
2r_\oplus$ and the $r\a < 2r_\oplus/\sqrt{2}$ regimes, and is plotted in
Fig.~\ref{fig:occ_rate_model_3_log}.

@@

We explore this quantitatively in Fig.~\ref{fig:frac_model_3}.
The lines in this figure are computed using the expressions given in the 
appendix for the number of detections coming from singles, primaries, and 
secondaries (Eqs.~\ref{eq:n_det_0},~\ref{eq:n_det_1}, and~\ref{eq:n_det_2}).
In the scenario that high-resolution imaging finds a system with $r\a > 
2r_\oplus$ in a binary, of order $\sim 3\times$ more detected planets orbit 
primaries than secondaries, for the $Z_2/Z_0=1$ case.
For the $Z_2/Z_0=0.5$ case, of order $\sim 7\times$ more detected planet 
planets orbit primaries than secondaries.
At $r\a < 2r_\oplus$, the ratios are $\sim 1.6 \times$ and $\sim 3.1 \times$ 
as many detected planets orbiting primaries than secondaries for the same 
$Z_2/Z_0$ values.

@@

In the recent work by~\citet{petigura_CKS_2017}, hot Jupiters appear as an 
island in period-radius space, rather than as a continuous component of a 
power law distribution.
Consider instead a Gaussian radius shape function,
\begin{equation}
f(r) \propto \exp \left( -\frac{(r-\bar{r})^2}{2\sigma^2} \right),
\end{equation}
with $\bar{r} = 14r_\oplus$ and $\sigma = 2r_\oplus$.
As always, $\Gamma_i(r) = Z_i f(r)$.
We then compute $\Gamma\a(r\a)$, and plot it in Fig.~\ref{fig:gaussian_HJ}.
Integrating to find hot Jupiter rates,
we find the opposite effect as in the power law model.
For $Z_2/Z_0>0$, 
the apparent HJ rate is {\it greater} than the true rate for singles.
The effect is maximal when there are as many hot Jupiters orbiting secondaries 
as singles, in which case
$\Lambda_{{\rm HJ},0}/\Lambda_{\rm HJ,a} = 0.84$
For the case when no hot Jupiters orbit secondaries, $\Lambda_{{\rm HJ},0} 
\approx \Lambda_{\rm HJ,a}$ to within one percent.

@@

\paragraph{Fraction of detected planets with binary companions vs. apparent 
radius}
A prediction of Fig.~\ref{fig:frac_model_3} is that the fraction of detected 
planets with binary companions increases by $\approx 6-12\%$ going from 
$2r_\oplus$ to $\approx 1.4r_\oplus$.
In \citet{ziegler_robo-ao_2017}'s recent summary of the Robo-AO KOI survey, 
they reported the nearby star fraction rates for planet candidates binned by 
Earths ($r\a < 1.6r_\oplus$), Neptunes ($1.6r_\oplus < r\a < 3.9 r_\oplus$), 
Saturns ($3.9r_\oplus < r\a < 9 r_\oplus$), and Jupiters ($r\a > 9r_\oplus$).
The nearby star rates and their $1\sigma$ uncertainties 
(following~\cite{burgasser_binarity_2003}'s procedure) are as
follows:
\begin{itemize}
    \item Earths: $16.3 \pm 1.0\%$, from 1480 systems.
    \item Neptunes: $13.0 \pm 0.8\%$, from 2058 systems.
    \item Saturns: $13.6 \pm 2.0\%$, from 338 systems.
    \item Jupiters: $19.0 \pm 2.8\%$, from 247 systems.
\end{itemize} 
where the values of the rates are less than the $\sim 45\%$ typical for 
Sun-like stars at least in part because of Robo-AO's sensitivity 
(\citealt{ziegler_robo-ao_2017}'s Fig.~2;~\citealt{raghavan_survey_2010}).
The uptick in the companion rate for Jupiters is likely tied to a large 
astrophysical false positive rate~\citep{santerne_sophie_2012}.
Going from Neptunes to Earths, the data suggest a weak increase in the 
detected planet companion fraction, perhaps of a few percent.

@@

Assuming a broken power-law radius distribution, we showed in 
Fig.~\ref{fig:frac_model_3} that assuming a detected planet has equal 
probability of orbiting the primary or secondary leads to an overestimate of
binarity's population-level effects.
A planet orbiting the secondary does lead to extreme corrections.
However, at $r\a \gtrsim 2r_\oplus$, these cases are rare outliers.
At $r\a \lesssim 2r_\oplus$, for all of the trial values of $Z_2/Z_0$ assumed 
in Fig.~\ref{fig:frac_model_3}, a detected planet is always more likely to 
orbit the primary than the secondary.

>>>>

We are mostly interested in the apparent rate density's radius dependence.
Marginalizing Eq.~\ref{eq:general_Gamma_a} over apparent stellar mass, we find
that when $Z_0=Z_1=Z_2$,
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma\a(r\a)}{\Gamma_0(r)} 
\right|_{r\a\rightarrow r} \\
&=
\notag
\frac{1}{1+\mu}
\left[1 + \frac{1}{\mathcal{N}_q} \frac{{\rm BF}}{1-{\rm BF}}\times 
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+4}{2}} +
\right.
\right. \\
&\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
\left.\left.
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{5}{3}} 
(1+q^\alpha)^{\frac{3}{2}}
(1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\label{eq:powerlaw_vary_binary}
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\gamma=0$, $\delta=-2.92$, the 
summed integrals in Eq.~\ref{eq:powerlaw_vary_binary} give $(\ldots)\approx 
1.50249$. %171211_model_integrals.nb
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma_0 = 1.048$; the
apparent rate density is higher than the rate density around singles by 4.8\%.
In summary, this model suggests that for a power-law radius distribution 
($2r_\oplus \lesssim r\a \lesssim 17r_\oplus$) in which there are the same 
number of planets orbiting singles, primaries, and secondaries, binarity 
influences apparent planet occurrence rates around Sun-like stars at the 
$\sim$few percent level.

@@

At apparent radii $r\a > 2r_\oplus$, the results of this model are the same as 
those from Sec.~\ref{sub:powerlaw_varying_binaries}.
For $r\a < 2r_\oplus$, the equations are quite tedious, though still tractable.
We insert Eq.~\ref{eq:model3_radius_distribution} into 
Eq.~\ref{eq:general_Gamma_a}, and integrate using a computer 
program. We refer the interested reader to our online 
implementation\footnote{\url{https://github.com/lgbouma/binary_biases}}.
The output is validated against analytic predictions in the $r\a > 
2r_\oplus$ and the $r\a < 2r_\oplus/\sqrt{2}$ regimes, and is plotted in
Fig.~\ref{fig:occ_rate_model_3_log}.

@@

We explore this quantitatively in Fig.~\ref{fig:frac_model_3}.
The lines in this figure are computed using the expressions given in the 
appendix for the number of detections coming from singles, primaries, and 
secondaries (Eqs.~\ref{eq:n_det_0},~\ref{eq:n_det_1}, and~\ref{eq:n_det_2}).

In short, the answer to our question is ``it depends on the number of planets 
orbiting secondaries, and also on the apparent radius of the detected planet''.
In the scenario that high-resolution imaging finds a system with $r\a > 
2r_\oplus$ in a binary, roughly $\sim 2.5\times$ more detected planets orbit 
primaries than secondaries, for the $Z_2/Z_0=1$ case.
For the $Z_2/Z_0=0.5$ case, $\sim 5\times$ more detected planet 
planets orbit primaries than secondaries.
So, for ``large'' apparent radii (relative to the cutoff in the 
intrinsic rate density), yes, detected planets in binaries are more 
likely to orbit the primary.

The situation for $r\a < 2r_\oplus$ is more nuanced.
For the $Z_2=0$ and $Z_2/Z_0=0.5$ cases, detected planets in binaries are 
still always more likely to orbit the primary.
However, if there are as many planets orbiting the secondaries as the 
primaries, there is actually a turn-over radius, $\approx 0.4r_\oplus$, 
beyond which more of the detected planets in binaries come from secondaries: 
they have all been ``diluted down'' from larger true planetary radii!
Although this (apparent) radius is at the limit of current sensitivities, this 
effect will be interesting for future instruments to investigate.

@@

In the recent work by~\citet{petigura_CKS_2017}, hot Jupiters appear as an 
island in period-radius space, rather than as a continuous component of a 
power law distribution.
Consider instead a Gaussian radius shape function,
\begin{equation}
f(r) \propto \exp \left( -\frac{(r-\bar{r})^2}{2\sigma^2} \right),
\end{equation}
with $\bar{r} = 14r_\oplus$ and $\sigma = 2r_\oplus$.
As always, $\Gamma_i(r) = Z_i f(r)$.
We then compute $\Gamma\a(r\a)$, and plot it in Fig.~\ref{fig:gaussian_HJ}.
Integrating to find hot Jupiter rates,
we find the opposite effect as in the power law model.
For $Z_2/Z_0>0$, 
the apparent HJ rate is {\it greater} than the true rate for singles.
The effect is maximal when there are as many hot Jupiters orbiting secondaries 
as singles, in which case
$\Lambda_{{\rm HJ},0}/\Lambda_{\rm HJ,a} = 0.81$.
%int_rate_density_vs_radius_model_7_withtext*manyZs.pdf
For the case when no hot Jupiters orbit secondaries, $\Lambda_{{\rm HJ},0} 
\approx \Lambda_{\rm HJ,a}$ to within one percent.

@@

\paragraph{Fraction of detected planets with binary companions vs. apparent 
radius}
A prediction of Fig.~\ref{fig:frac_model_3} is that the fraction of detected 
planets with binary companions increases by $\approx 6-12\%$ going from 
$2r_\oplus$ to $\approx 1.4r_\oplus$.
In \citet{ziegler_robo-ao_2017}'s recent summary of the Robo-AO KOI survey, 
they reported the fraction of planet-hosting stars with Robo-AO detected 
companion stars, binning by Earths ($r\a < 1.6r_\oplus$), Neptunes 
($1.6r_\oplus < r\a < 3.9 r_\oplus$), Saturns ($3.9r_\oplus < r\a < 9 
r_\oplus$), and Jupiters ($r\a > 9r_\oplus$).
The reported nearby star rates and their $1\sigma$ uncertainties are:
\begin{itemize}
    \item Earths: $16.3 \pm 1.0\%$, from 1480 systems.
    \item Neptunes: $13.0 \pm 0.8\%$, from 2058 systems.
    \item Saturns: $13.6 \pm 2.0\%$, from 338 systems.
    \item Jupiters: $19.0 \pm 2.8\%$, from 247 systems.
\end{itemize} 
where the uncertainties were calculated 
following~\citet{burgasser_binarity_2003}.
The absolute values of the rates are less than the $\sim 45\%$ typical for 
Sun-like stars at least in part because of Robo-AO's sensitivity 
(\citealt{ziegler_robo-ao_2017}'s Fig.~2;~\citealt{raghavan_survey_2010}).
Another reason for the lower measured companion rates could be that planetary 
systems are less likely to have 
binary companions (as~\citealt{kraus_impact_2016} argued for close binary 
companions).
The uptick in the companion rate for Jupiters is likely tied to a large 
astrophysical false positive rate~\citep{santerne_sophie_2012}.
Going from Neptunes to Earths, the data suggest a weak increase in the 
detected planet companion fraction, perhaps of a few percent.

@@

Assuming a broken power-law radius distribution, we showed in 
Fig.~\ref{fig:frac_model_3} that assuming a detected planet has equal 
probability of orbiting the primary or secondary probably is rarely correct.
A planet orbiting the secondary does lead to extreme corrections.
However, at $r\a \gtrsim 2r_\oplus$, these cases are rare outliers.
At $r_\oplus \lesssim r\a \lesssim 2r_\oplus$, for all of the trial values of 
$Z_2/Z_0$ assumed in Fig.~\ref{fig:frac_model_3}, a detected planet is always 
more likely to orbit the primary than the secondary.
Only at very small apparent radii, $r\a \approx 0.5r_\oplus$, and only if 
secondaries host as many planets as singles and primaries, does this norm 
break down.
